Carbonate chemistry at Ny-Ålesund

### 2021-05-21 awipev-CO2.Rmd
added code from Kapsenberg's paper

@@

>>>>

# Same plot for paper
p_at_s <- ggplot(d_all_at_discrete,aes(x=sal_mix, y= at)) +
  geom_point(col="blue", size = 0.2) +
  labs(x="Salinity", y="Discrete AT")  +
  stat_smooth(method = "lm", col = "blue", size = 0.3) +
  theme_bw(base_size = 9) +
  theme(aspect.ratio=1#,
        #axis.text=element_text(size=7),
        #axis.title.y= element_text(size =7),
        #axis.title.x= element_text(size =7),
        #plot.margin = unit(c(0.5,1,1,1), "cm"),
        #plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70))
        )
ggsave(filename = "figures/p_at_s.png", plot = p_at_s, width = 7, height = 7, units = "cm")

@@

## Deconvolution of pCO2 changes
```{r pCO2 changes due to changes in T, S, AT and CT, include=FALSE}
# We follow the method of Garcia-Ibanez described above.
# construct a dataframe with monthly means for the entire series (mXX) as well as real data (without mrefix m)
deconv_dat <- tbl_df(
  data.frame(date=as.POSIXct(mean0m$sampling_date), mAT=means_month0m$AT[mean0m$monthd]*1e-6, mCT=means_month0m$CT[mean0m$monthd]*1e-6, mS=means_month0m$Salinity[mean0m$monthd], mT=means_month0m$Temperature[mean0m$monthd], P=mean0m$mean_P, Pt=mean0m$mean_Pt, Sit=mean0m$mean_Sit, AT=mean0m$mean_ta*1e-6, CT=mean0m$mean_dic*1e-6, S=mean0m$mean_salinity, T=mean0m$mean_temperature)
)

d <- NULL; deconv <- NULL # d is for the data to use in carb, deconv is to store the results
deconv <- data_frame(date=as.POSIXct(mean0m$sampling_date))
d <- filter(deconv_dat, !is.na(mAT), !is.na(mCT), !is.na(mS), !is.na(T)) # we eliminate date with NAs
deconv_pCO2_T <- data_frame(date=d$date,
                     deconv_pCO2_T=carb(flag=15, var1=d$mAT, var2=d$mCT, S=d$mS, T=d$T, P=d$P, Pt=d$Pt, Sit=d$Sit, k1k2 = "l", kf = "pf", pHscale = "T", b="l10")$pCO2
)
d <- filter(deconv_dat, !is.na(mAT), !is.na(mCT), !is.na(S), !is.na(mT)) # we eliminate date with NAs
deconv_pCO2_S <- data_frame(date=d$date,
                     deconv_pCO2_S=carb(flag=15, var1=d$mAT, var2=d$mCT, S=d$S, T=d$mT, P=d$P, Pt=d$Pt, 
                                      Sit=d$Sit, k1k2 = "l", kf = "pf", pHscale = "T", b="l10")$pCO2
                     )
d <- filter(deconv_dat, !is.na(AT), !is.na(mCT), !is.na(mS), !is.na(mT)) # we eliminate date with NAs
deconv_pCO2_AT <- data_frame(date=d$date,
                     deconv_pCO2_AT=carb(flag=15, var1=d$AT, var2=d$mCT, S=d$mS, T=d$mT, P=d$P, Pt=d$Pt, Sit=d$Sit, k1k2 = "l", kf = "pf", pHscale = "T", b="l10")$pCO2
)
d <- filter(deconv_dat, !is.na(mAT), !is.na(CT), !is.na(mS), !is.na(mT)) # we eliminate date with NAs
deconv_pCO2_CT <- data_frame(date=d$date,
                     deconv_pCO2_CT=carb(flag=15, var1=d$mAT, var2=d$CT, S=d$mS, T=d$mT, P=d$P, Pt=d$Pt, Sit=d$Sit, k1k2 = "l", kf = "pf", pHscale = "T", b="l10")$pCO2
)

# Bring all results together
deconv <- deconv %>%
  left_join(deconv_pCO2_T) %>%
  left_join(deconv_pCO2_S) %>%
  left_join(deconv_pCO2_AT) %>%
  left_join(deconv_pCO2_CT)

# Regressions and tables of key parameters
var_list <-
  c(
  "deconv_pCO2_T",
  "deconv_pCO2_S",
  "deconv_pCO2_AT",
  "deconv_pCO2_CT"
  )
lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(date), list(i = as.name(x))), data = deconv))
  })
reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
  pf(lms[[i]]$fstatistic[1],
  lms[[i]]$fstatistic[2],
  lms[[i]]$fstatistic[3],
  lower.tail = FALSE)
  reg <-
  rbind(reg, as.numeric(
  c(
    lms[[i]]$coefficients[2, 1],
    lms[[i]]$coefficients[2, 2],
  lms[[i]]$coefficients[2, 4],
  lms[[i]]$coefficients[1, 1],
  lms[[i]]$coefficients[1, 2],
  lms[[i]]$coefficients[1, 4],
  lms[[i]]$r.squared,
  prob
  )
  ))
}
colnames(reg) <-
  c("Slope",
  "SE Slope",
  "P Slope",
  "Intercept",
  "SE int.",
  "P int.",
  "R2",
  "P value")
row.names(reg) <- var_list
reg_deconv <- reg

# Plot
#deconv_melt <- melt(deconv, id="date")
deconv_melt <-  gather(data = deconv, key = variable, value=value, -date)
fig_deconv_pCO2 <- ggplot(deconv_melt) + geom_point(aes(x=date, y=value, color=variable), size=0.5) +
  facet_grid(variable ~., scales="free_y") + 
  xlab("Time") + ylab("pCO2 (uatm")+
  Mytheme() +
  scale_colour_discrete(guide="none")
ggsave(file="../figures/fig_deconv_pCO2.png", fig_deconv, width = 18, height = 20, units = "cm")
```
